# Feature selection
# Yunlong Jiao, 02 May 2016

This script studies feature selection with representative FS algorithms to distinguish importance of features. And finally sees which feature type when a model is learned based on multiple feature types is more interesting in an overall comparison. Specifically we
- run independent significance t.test gene-by-gene or pathway-by-pathway and see the p-values from different types of features
- look at feature selection results from prediction made with pulled features of path.vals and genes.vals

```{r setup, message = FALSE}
knitr::opts_chunk$set(error = FALSE, fig.width = 12, fig.height = 8, dev = "pdf", fig.keep = "high", fig.path = "5_featselect_figure/", cache.path = "5_featselect_cache/")
set.seed(17766220)
source("../../src/func.R")
library(reshape2)
library(ggplot2)
```

First load in workspace and read in parameters!

```{r param}
# load in workspace
load("dat.RData")
rm(list = ls(pattern="[.]kmat$"))
# read in parameters
param <- read.table("2_runPredict.txt", header = FALSE, row.names = NULL, col.names = c("xname", "yname", "prname", "i.fold", "nfolds", "nrepeats", "i.fold.inn", "nfolds.inn", "nrepeats.inn"))
# features
xlist <- unique(param$xname)
xlist
# feature types # the following order is fairly important
xlist.type <- c("func-wise", "func-wise", 
                "path-wise", "path-wise", 
                "gene-wise", "gene-wise", "gene-wise", 
                "mix-wise", "mix-wise", "mix-wise")
names(xlist.type) <- c("fun.vals", "go.vals", # functionality features
                       "eff.vals", "path.vals", # pathway features
                       "mini.genes.vals", "other.genes.vals", "genes.vals", # gene features
                       "eff.and.other.genes.vals", "path.and.other.genes.vals", "path.and.genes.vals") # mixed type
xlist.vline <- c(2.5, 4.5, 7.5) # cut out types
stopifnot(length(setdiff(xlist, names(xlist.type))) == 0)
xlist.type
# features to perform indep signif test
xlist.test <- grep("[.]and[.]|^genes", xlist, value = TRUE, invert = TRUE)
xlist.test <- xlist.test[order(match(xlist.test, names(xlist.type)))]
xlist.test
# groups
ylist <- unique(param$yname)
ylist
# predictors
prlist <- unique(param$prname)
prlist
# predictors that automate feat select
prlist.fs <- c("predictorLogitLasso","predictorPAM","predictorRF")
prlist.fs <- prlist.fs[prlist.fs %in% prlist]
prlist.fs
# (outter) `nfolds`-fold CV repeated `nrepeats` times for evaluation
nfolds <- unique(param$nfolds)
stopifnot(length(nfolds) == 1)
nfolds
nrepeats <- unique(param$nrepeats)
stopifnot(length(nrepeats) == 1)
nrepeats
# inner `nfolds.inn`-fold CV repeated `nrepeats.inn` times for tuning predictors
nfolds.inn <- unique(param$nfolds.inn)
stopifnot(length(nfolds.inn) == 1)
nfolds.inn
nrepeats.inn <- unique(param$nrepeats.inn)
stopifnot(length(nrepeats.inn) == 1)
nrepeats.inn
# evaluation measures
slist <- c("acc","fpr","tpr","ppv","fval", "auroc")
slist
slist.prefer.large.score <- c(TRUE, FALSE, TRUE, TRUE, TRUE, TRUE)
names(slist.prefer.large.score) <- slist
slist.prefer.large.score
```

## Independent significance t.test

For each `yname` in `r ylist`, we run independent significance t.test for each pathway and each gene individually and look at the returned p-values. (In case of multiple levels, we run t.test against binary labels that discriminate the last level vs the other levels.) In the end we have several different plots to illustrate the results.

```{r test_prepare, cache=TRUE, message=FALSE}
# read in param
param2 <- read.table("4_indepsignif.txt", header = FALSE, row.names = NULL, col.names = c("xname", "yname", "i.fold", "nfolds", "nrepeats", "pthres", "test", "method"))
stopifnot(length(setdiff(xlist.test, unique(param2$xname))) == 0)
stopifnot(length(setdiff(ylist, unique(param2$yname))) == 0)
stopifnot(all(nfolds == unique(param2$nfolds)))
stopifnot(all(nrepeats == unique(param2$nrepeats)))
# pthres
pthres <- unique(param2$pthres)
stopifnot(length(pthres) == 1)
pthres
# test
test <- unique(param2$test)
stopifnot(length(test) == 1)
test
# method for multiple correction
method <- unique(param2$method)
stopifnot(length(method) == 1)
method
# show each grp separately
for (yname in ylist) {
  plist <- lapply(xlist.test, function(xname){
    message(yname, '\t', xname)
    res.files <- list.files(path = 'Robj', 
                            pattern = paste('^plist', xname, yname, 
                                            '[[:digit:]]+', nfolds, nrepeats, 
                                            pthres, test, method, 
                                            sep = '_'), 
                            full.names = TRUE)
    res <- lapply(res.files, function(f){
      v <- get(load(f))
      stopifnot(is.vector(v))
      as.data.frame(t(v)) # convert to data.frame to facilitate further rbind
    })
    res <- do.call('rbind', res)
    return(colMeans(res)) # averaged p-value over CV splits of data
  })
  names(plist) <- xlist.test
  assign(paste0(yname,"_plist"), plist)
}
```

For each `yname` in `r ylist`, we have several plots to illustrate the results
- Boxplots of p-values of feature-by-feature t.test is used to show the variance of feature significance for different feature types. Two plots correpond to that with all features, or that with only significant features at a p-value less than `pthres`.
- Lineplot of the accumulated proportion defined by the number of significant features within each type (at increasing p-value thresholds) divided by the total number of features from that specific type. The number on line is the absolute count of number of significant features within each type.
- Barplot of the proportion defined by the number of significant features within each type divided by the number of significant features of all types pulled together (at increasing p-value thresholds). The number on bar is the absolute count of number of significant features within each type.

```{r test, results='hold'}
cat('\n---> \t Independent significance', test, '\t <---\n')
# show each grp separately
for (yname in ylist) {
  cat('\n---------> \t for ', yname, ' \t <---------\n')
  plist <- get(paste0(yname,"_plist"))
  # show top 10 signif features
  print(lapply(plist, head, n = 10))
  plist.melt <- melt(plist, level = 1, value.name = "value")
  colnames(plist.melt)[colnames(plist.melt)=="L1"] <- "xname"
  plist.melt$xname <- factor(plist.melt$xname, levels = xlist.test, ordered = TRUE)
  
  # feature-by-feature p-values
  cat('\nBoxplots of p-values of feature-by-feature', test, '\n')
  p1 <- ggplot(plist.melt, aes(x = xname, y = value)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(trans = "log10", limits = c(1/1e6, 1), name = "p-value") + 
    ggtitle("All features") + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  p2 <- ggplot(subset(plist.melt, value < pthres), aes(x = xname, y = value)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(trans = "log10", limits = c(pthres/1e6, pthres), name = "p-value") + 
    ggtitle(paste0("Significant features only (p-value<", pthres, ")")) + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  multiplot(p1, p2, cols = 2)
  
  # make a list of pthres in log-scale and count the number of features being signif within each type
  pthres.min <- quantile(plist.melt$value[plist.melt$value<pthres], 0.01)
  pthreslist <- pthres.min*exp(seq(0,1,length.out=20)*log(pthres/pthres.min))
  plist.tab <- lapply(pthreslist, function(pthres){
    table(subset(plist.melt, subset = (value < pthres), select = "xname"))
  })
  names(plist.tab) <- formatC(pthreslist, digits = 2, format = "e")
  plist.tab <- do.call('rbind', plist.tab)
  
  # compared within each type
  plist.freq <- melt(plist.tab, varnames = c("pthres", "type"), value.name = "value")
  n.featlist <- sapply(xlist.test, function(xname) ncol(get(xname)))
  plist.freq$freq <- plist.freq$value/n.featlist[plist.freq$type]
  cat('\nTotal number of features within each type\n')
  print(n.featlist)
  p1 <- ggplot(plist.freq, aes(x = factor(pthres), y = freq)) + 
    geom_line(aes(group = type, colour = type), stat="identity", alpha = 0.8) + 
    geom_text(aes(label = value, colour = type)) + 
    scale_x_discrete(name = "at p-value threshold (less than)") + 
    scale_y_continuous(name = "Proportion (%)") + 
    ggtitle("Accumulated proportion of signif features within each type (%)")
  plot(p1)
  
  # pull and compare across types
  plist.pull <- melt(plist.tab, varnames = c("pthres", "type"), value.name = "value")
  n.featsignif <- rowSums(plist.tab)
  plist.pull$freq <- plist.pull$value/n.featsignif[match(plist.pull$pthres, as.numeric(names(n.featsignif)))]
  p1 <- ggplot(plist.pull, aes(x = factor(pthres), y = freq)) + 
    geom_bar(aes(fill = type), stat="identity", position = "stack", alpha = 0.8) + 
    geom_text(aes(label = value), colour = "black", position = "stack", vjust = 0.9) + 
    scale_x_discrete(name = "at p-value threshold (less than)") + 
    scale_y_continuous(name = "Proportion (%)") + 
    ggtitle("Accumulated proportion of signif features across different types (%)")
  plot(p1)
}
```

## Algorithmic feature selection

For each `yname` in `r ylist`, prediction is made with pulled types of features, that is a naive combination of path.vals and genes.vals. We look at the frequency of each type of features being selected automatically.

```{r featselect_prepare, cache=TRUE, message=FALSE}
xname <- "path.and.genes.vals" # do not alter
xname.list <- c("mini.genes.vals", "other.genes.vals", "path.vals") # do not alter
featlist.long <- lapply(xname.list, function(u){
  colnames(get(u))
})
names(featlist.long) <- xname.list
n.featlist.long <- sapply(featlist.long, length)
# show each grp separately
for (yname in ylist) {
  for (prname in prlist.fs) {
    message(yname, '\t', prname)
    res.files <- list.files(path = 'Robj', 
                            pattern = paste('^ivres', xname, yname, prname, 
                                            '[[:digit:]]+', nfolds, nrepeats, 
                                            0, nfolds.inn, nrepeats.inn, 
                                            sep = '_'), 
                            full.names = TRUE)
    res <- lapply(res.files, function(f) get(load(f)))
    names(res) <- paste0("rep",1:length(res))
    
    # scores
    scores <- lapply(res, '[', slist)
    scores <- melt(scores, level = 1, value.name = "value")
    colnames(scores)[colnames(scores)=="L1"] <- "rep"
    colnames(scores)[colnames(scores)=="L2"] <- "score"
    scores$prname <- prname
    rownames(scores) <- 1:nrow(scores)
    assign(paste(yname,prname,"scores",sep="_"), scores)
    
    # featselect
    featlist.short <- lapply(res, '[[', 'featlist.short')
    featlist.short <- lapply(featlist.long, function(u){
      lapply(featlist.short, function(v){
        intersect(u,v)
      })
    })
    assign(paste(yname,prname,"featlist.short",sep="_"), featlist.short)
    rm(res)
  }
}
```

For each `yname` in `r ylist`, we have several plots to illustrate the results
- Boxplot of prediction performance using pulled feature set `r xname`, wrt different criterion and different prediction algorithm, where variance shows over `r nfolds*nrepeats` CV runs.
- Three boxplots on the number of selected features from each feature type in `r xname.list`, wrt different algorithms, where variance shows at each CV run. Three of them each correpond to 1) total count of selected features from each type; 2) total count of selected features from each type divided by the total number of selected features of all types; 3) total count of selected features from each type divided by the total number of features from that same type.
- Histogram of how many times one feature are selected in total over all `r nfolds*nrepeats` CV runs, specific to each type and each algorithm.

```{r featselect, results='hold'}
cat('\n---> \t Algorithmic feature selection \t <---\n')
# show each grp separately
for (yname in ylist) {
  cat('\n---------> \t for ', yname, ' \t <---------\n')
  # scores
  scores <- lapply(prlist.fs, function(prname){
    get(paste(yname,prname,"scores",sep="_"))
  })
  scores <- do.call('rbind', scores)
  scores$value[scores$score == "fpr"] <- 1 - scores$value[scores$score == "fpr"]
  scores$score[scores$score == "fpr"] <- "spec"
  scores$score[scores$score == "tpr"] <- "sens"
  p1 <- ggplot(scores, aes(x = prname, y = value)) + 
    geom_boxplot(aes(fill = prname), alpha = 0.8) + 
    facet_wrap(~score, scales = "free") + 
    ggtitle(paste0("predicting ", yname, " with ", xname)) + 
    theme(axis.text.x = element_blank())
  plot(p1)
  
  # featselect
  featlist.short <- lapply(prlist.fs, function(prname){
    get(paste(yname,prname,"featlist.short",sep="_"))
  })
  names(featlist.short) <- prlist.fs
  # featlist.short is 3-layer list prlist.fs -> xname.list -> CV repeats -> a character vector of feature names
  # number of features are selected at each CV run
  featlist.freq <- lapply(featlist.short, function(u) lapply(u, lapply, length))
  featlist.freq <- melt(featlist.freq, level = 1, value.name = "value")
  colnames(featlist.freq)[colnames(featlist.freq)=="L1"] <- "prname"
  colnames(featlist.freq)[colnames(featlist.freq)=="L2"] <- "xname"
  colnames(featlist.freq)[colnames(featlist.freq)=="L3"] <- "rep"
  p1 <- ggplot(featlist.freq, aes(x = xname, y = value)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    facet_wrap(~prname, scales = "free") + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(name = "Number of selected features") + 
    ggtitle(paste0("How many features are selected at each run")) + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  plot(p1)
  # proportion across group
  n.featlist.short <- tapply(featlist.freq$value, list(featlist.freq$rep, featlist.freq$prname), sum)
  cat('\nTotal number of features selected at each run (averaged over each CV run)\n')
  print(colMeans(n.featlist.short))
  featlist.freq$prop <- featlist.freq$value/n.featlist.short[cbind(featlist.freq$rep,featlist.freq$prname)]
  p2 <- ggplot(featlist.freq, aes(x = xname, y = prop)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    facet_wrap(~prname, scales = "free") + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(name = "Proportion") + 
    ggtitle(paste0("Proportion of features selected across types at each run")) + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  plot(p2)
  # proportion within group
  cat('\nTotal number of features within each type\n')
  print(n.featlist.long)
  featlist.freq$freq <- featlist.freq$value/n.featlist.long[featlist.freq$xname]
  p3 <- ggplot(featlist.freq, aes(x = xname, y = freq)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    facet_wrap(~prname, scales = "free") + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(name = "Proportion") + 
    ggtitle(paste0("Proportion of features selected within own type at each run")) + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  plot(p3)
  
  # show top 3 most selected features in each type by predictor
  featlist.tab <- lapply(featlist.short, function(u) lapply(u, function(v) 
    sort(table(unlist(v)), decreasing = TRUE)
  ))
  cat('\nPreview of top 3 most often selected features in each type\n')
  print(lapply(featlist.tab, function(u) lapply(u, head, n=3)))
  
  # count number of times of features from each type being selected over CV runs
  featlist.count <- melt(featlist.tab, level = 1, value.name = "value")
  colnames(featlist.count)[colnames(featlist.count)=="L1"] <- "prname"
  colnames(featlist.count)[colnames(featlist.count)=="L2"] <- "xname"
  colnames(featlist.count)[colnames(featlist.count)=="Var1"] <- "featname"
  featlist.count <- dcast(featlist.count, prname+xname~value, fun.aggregate = length)
  featlist.count <- melt(featlist.count, id.vars = c("prname","xname"))
  p1 <- ggplot(featlist.count, aes(x = variable, y = value)) + 
    geom_bar(aes(fill = xname), stat = "identity", position = "dodge", alpha = 0.8) + 
    coord_flip() + 
    geom_text(aes(label = value, group = xname), colour = "black", position = position_dodge(0.9), vjust = 0.9) + 
    facet_wrap(~prname) + 
    scale_x_discrete(name = paste0("selected times (max ", nfolds*nrepeats, ")")) + 
    scale_y_continuous(name = "count") + 
    ggtitle(paste0("How many times features are selected over all ", nfolds*nrepeats, " CV runs")) + 
    theme(legend.title = element_blank(), legend.position = "bottom")
  plot(p1)
}
```

## Session info

```{r session_info}
sessionInfo()
```
