# Process ICGC BRCA data
# Yunlong Jiao, 10 March 2016

This script processes data and generates matrices for naive models later to make predictions.

```{r setup}
knitr::opts_chunk$set(error = FALSE)
set.seed(35875954)
source("../../src/func.R")
datapath <- '../../data/BRCA_saved_data/'
```

# Features

Feature matrices are stored at `crom01:/fsclinic/common/analisis/data/BRCA_saved_data/`. Note that all feature matrices have to be initially provided by features in rows and patients in cols (to facilitate processing in a universal style). These matrix objects should have a name ending with `[.]vals$`.

```{r features}
# gene expression
load(paste0(datapath, '110_genes_vals.RData'))

# gene expression wrt those present in pathways
load(paste0(datapath, '110_mini_genes_vals.RData'))

# pathway-wise features
load(paste0(datapath, '110_hipathia_matrices.RData'))

# complement of mini.genes from all.genes
stopifnot(rownames(mini.genes.vals) %in% rownames(genes.vals))
other.genes.vals <- genes.vals[setdiff(rownames(genes.vals), rownames(mini.genes.vals)), ]

# show all feature matrices
xlist <- ls(pattern = '[.]vals$')
xlist
for (xname in xlist) {
  # process each feature matrices
  x <- get(xname)
  # check no NA values
  stopifnot(!any(is.na(x)))
  # make patients in rows and features in cols
  x <- t(x)
  # reform feature names and check for duplicated names
  colnames(x) <- paste0('X_', gsub('[^[:alnum:]_]', '_', colnames(x)))
  stopifnot(!any(duplicated(rownames(x))))
  stopifnot(!any(duplicated(colnames(x))))
  # assign back
  assign(xname, x)
  # preview
  cat('\n-------------> ', xname, ' <-------------\n')
  print(str(get(xname)))
  cat('\n')
}
# set samples
samplelist <- unique(lapply(xlist, function(xname) rownames(get(xname))))
stopifnot(length(samplelist) == 1)
samplelist <- unlist(samplelist)
nsample <- length(samplelist)
```

# Groups (binary)

Binary classes are created from clinical info for classification. There vector objects are named ending with `[.]grps$`.

```{r groups}
# basal type
y <- read.table(paste0(datapath, 'sampleBasalType.txt'), header = FALSE)
head(y$V2)
basal.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$V2), levels = c("Not_Basal", "Basal"), labels = c("neg_NonBasal", "pos_Basal"))
stopifnot(!any(is.na(basal.grps)))
names(basal.grps) <- y$V1
head(basal.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(basal.grps)))
table(basal.grps[sl])

# phenotype tumor
y <- read.table(paste0(datapath, 'TCGA_phenotype.txt'), header = TRUE)
head(y$cancer)
tumor.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$cancer), levels = c("Normal", "Tumor"), labels = c("neg_Normal", "pos_Tumor"))
stopifnot(!any(is.na(tumor.grps)))
names(tumor.grps) <- rownames(y)
head(tumor.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(tumor.grps)))
table(tumor.grps[sl])

# survival
alignmat <- read.table(paste0(datapath, 'clinical_info_sample.txt'), sep = "\t", quote = "\"", header = TRUE)
y <- read.table(paste0(datapath, 'clinical_info_donor.txt'), sep = "\t", quote = "\"", header = TRUE)
alignid <- data.frame(sample = samplelist, 
                      donor = sapply(samplelist, function(id){
                        i <- grep(id, alignmat$submitted_sample_id)
                        if (length(i) == 1) return(alignmat$icgc_donor_id[i])
                        else if (length(i) == 0) return(NA)
                        else stop(id, " has more than 1 alignment!")
                      }))
# number of samples before pruning
nrow(alignid)
alignid <- subset(alignid, !is.na(alignid$donor))
id <- match(alignid$donor, y$icgc_donor_id)
alignid <- alignid[!is.na(id), ]
id <- id[!is.na(id)]
y <- y[id, ]
# number of samples after pruning
nrow(alignid)
head(y$donor_vital_status)
surv.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$donor_vital_status), levels = c("alive", "deceased"), labels = c("neg_alive", "pos_deceased"))
stopifnot(!any(is.na(surv.grps)))
names(surv.grps) <- alignid$sample
head(surv.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(surv.grps)))
table(surv.grps[sl])

# show all label vectors
ylist <- ls(pattern = '[.]grps$')
ylist
# check for now to deal with binary classification only
for (yname in ylist) {
  stopifnot(length(levels(get(yname))) == 2)
}
```

# Predictors

See `../../src/func.R` for all predictors implemented. These function objects should all have a name starting with `^predictor`.

```{r predictors}
prlist <- ls(pattern = '^predictor')
prlist
```

# Save up !!

```{r save}
# write out combinations of datasets for running on cluster
param <- expand.grid(xlist, ylist, prlist, KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE)
write.table(param, file = 'cluster_param.txt', quote = FALSE, row.names = TRUE, col.names = FALSE)

# save entire image to be loaded later
save(list = c(xlist, ylist), file = 'dat.RData')
```

# Session info

```{r session_info}
sessionInfo()
```
