# Process ICGC BRCA data
# Yunlong Jiao, 10 March 2016

This script processes data and generates matrices for naive models later to make predictions.

```{r setup}
knitr::opts_chunk$set(error = FALSE, cache.path = "1_processData_cache/")
set.seed(35875954)
source("../../src/func.R")
datapath <- '../../data/BRCA_saved_data/'
```

## Features

Feature matrices are stored at `r datapath`. Note that all feature matrices have to be initially provided by features in rows and patients in cols (to facilitate processing in a universal style). These matrix objects should have a name ending with `[.]vals$`.

```{r features}
# gene expression
load(paste0(datapath, '110_genes_vals.RData'))

# gene expression wrt those present in pathways
load(paste0(datapath, '110_mini_genes_vals.RData'))

# pathway-wise features
load(paste0(datapath, '110_hipathia_matrices.RData'))

# complement of mini.genes from all.genes
stopifnot(rownames(mini.genes.vals) %in% rownames(genes.vals))
other.genes.vals <- genes.vals[setdiff(rownames(genes.vals), rownames(mini.genes.vals)), ]

# get all feature matrices
xlist <- ls(pattern = '[.]vals$')
for (xname in xlist) {
  # process each feature matrices
  x <- get(xname)
  # check no NA values
  stopifnot(!any(is.na(x)))
  # make patients in rows and features in cols
  x <- t(x)
  # reform feature names and check for duplicated names
  colnames(x) <- paste0('X_', gsub('[^[:alnum:]_]', '_', colnames(x)))
  stopifnot(!any(duplicated(rownames(x))))
  stopifnot(!any(duplicated(colnames(x))))
  # assign back
  assign(xname, x)
}
# set samples
samplelist <- unique(lapply(xlist, function(xname) rownames(get(xname))))
stopifnot(length(samplelist) == 1)
samplelist <- unlist(samplelist)
nsample <- length(samplelist)

# create mixed feature matrices

xlist2combn <- list("eff.and.other.genes.vals" = c("eff.vals", "other.genes.vals"), 
                    "eff.and.genes.vals" = c("eff.vals", "genes.vals"), 
                    "path.and.other.genes.vals" = c("path.vals", "other.genes.vals"), 
                    "path.and.genes.vals" = c("path.vals", "genes.vals"))
for (xname in names(xlist2combn)) {
  x <- mget(xlist2combn[[xname]])
  x <- do.call("cbind", x)
  assign(xname, x)
}
xlist <- c(xlist, names(xlist2combn))

# show all feature matrices
xlist
cat('\n-------------> \t preview \t <-------------\n')
for (xname in xlist) {
  cat('\n-------------> \t ', xname, ' \t <-------------\n')
  print(str(get(xname)))
  cat('\n')
}
```

## Groups

Classes labels are created from clinical info for classification. There vector objects are named ending with `[.]grps$`.

**NOTE for labels involving multiple classes, always put THE positive group to the last level in the ordered factor as the evaluation measure such as TPR, FPR will be specific to that level.**

#### Basal vs Non-basal (breast cancer basal type)

```{r basal, eval=FALSE}
y <- read.table(paste0(datapath, 'sampleBasalType.txt'), header = FALSE)
head(y$V2)
basal.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$V2), levels = c("Not_Basal", "Basal"), labels = c("neg_NonBasal", "pos_Basal"))
stopifnot(!any(is.na(basal.grps)))
names(basal.grps) <- y$V1
head(basal.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(basal.grps)))
# contrasting class size
table(basal.grps[sl])
```

#### Tumor vs Normal (tumor detection)

```{r tumor, eval=FALSE}
y <- read.table(paste0(datapath, 'TCGA_phenotype.txt'), header = TRUE)
head(y$cancer)
tumor.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$cancer), levels = c("Normal", "Tumor"), labels = c("neg_Normal", "pos_Tumor"))
stopifnot(!any(is.na(tumor.grps)))
names(tumor.grps) <- rownames(y)
head(tumor.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(tumor.grps)))
# contrasting class size
table(tumor.grps[sl])
```

#### Alive vs Deceased (overall survival)

```{r surv, message=FALSE, cache=TRUE}
alignmat <- read.table(paste0(datapath, 'clinical_info_sample.txt'), sep = "\t", quote = "\"", header = TRUE)
y <- read.table(paste0(datapath, 'clinical_info_donor.txt'), sep = "\t", quote = "\"", header = TRUE)
alignid <- data.frame(sample = samplelist, 
                      donor = sapply(samplelist, function(id){
                        message(".", appendLF = FALSE)
                        i <- grep(id, alignmat$submitted_sample_id)
                        if (length(i) == 1) {
                          return(alignmat$icgc_donor_id[i])
                        } else if (length(i) == 0) {
                          return(NA)
                        } else {
                          stop(id, " has more than 1 alignment!")
                        }
                      }))
# number of samples before pruning
nrow(alignid)
alignid <- subset(alignid, !is.na(alignid$donor))
id <- match(alignid$donor, y$icgc_donor_id)
alignid <- alignid[!is.na(id), ]
id <- id[!is.na(id)]
y <- y[id, ]
# number of samples after pruning
nrow(alignid)
head(y$donor_vital_status)
surv.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$donor_vital_status), levels = c("alive", "deceased"), labels = c("neg_alive", "pos_deceased"))
stopifnot(!any(is.na(surv.grps)))
names(surv.grps) <- alignid$sample
head(surv.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(surv.grps)))
table(surv.grps[sl])
```

#### Luminal vs Her2 vs Basal (breast cancer subtypes)

```{r subtype}
y <- read.table(paste0(datapath, 'sampleCancerType.txt'), header = FALSE)
head(y$V2)
unique(y$V2)
subtype.grps <- ordered(gsub('[^[:alnum:]_]', '_', y$V2), levels = c("Luminal", "Her2", "Basal"), labels = c("T1_Luminal", "T2_Her2", "T3_Basal"))
id.na <- is.na(subtype.grps)
subtype.grps <- subtype.grps[!id.na]
names(subtype.grps) <- y$V1[!id.na]
head(subtype.grps)
# sample size (how many labels are available in the sample cohort)
length(sl <- intersect(samplelist, names(subtype.grps)))
# contrasting class size
table(subtype.grps[sl])
```

#### Summary

```{r groups}
# get all groups
ylist <- ls(pattern = '[.]grps$')
# show all groups with their classes 
for (yname in ylist) {
  cat('\n-------------> \t summary \t <-------------\n')
  cat('\n-------------> \t ', yname, ' \t <-------------\n')
  print(levels(get(yname)))
  cat('\n')
}
```

## Predictors

See `../../src/func.R` for all predictors implemented. These function objects should all have a name starting with `^predictor`.

```{r predictors}
prlist <- ls(pattern = '^predictor.')
prlist
```

## Kernel matrices

This section computes kernel matrix to accelerate implementation of `kernlab::ksvm`. These kernel matrices are save at `r datapath` and named ending with `[.]kmat$`. Specifically we have the following kernel
- Kendall kernel (Jiao and Vert, 2016) for `r grep("kendall", prlist, ignore.case=T, value=T)`

**NOTE before we compute kernel here **

```{r kmat, message=FALSE}
# kendall kernel
prname <- "predictorKendallSVM"
kernel <- pcaPP::cor.fk
for (xname in xlist) {
  kmatname <- paste(prname, xname, "kmat", sep = ".")
  message(kmatname)
  kmatpath <- paste0(datapath, kmatname, ".RData")
  if (file.exists(kmatpath)) {
    assign(kmatname, get(load(kmatpath)))
  } else {
    x <- get(xname)
    x <- removeConst(x)
    assign(kmatname, computeKernelMatrix(x = x, kernel = kernel))
    save(list = kmatname, file = kmatpath)
  }
}
kmatlist <- ls(pattern = '[.]kmat$')
kmatlist
```

## (Nested) cross validation parameters

```{r cv}
# CV folds for evaluation
nfolds <- 5
nrepeats <- 10
# inner CV folds for tuning predictor
nfolds.inn <- 5
nrepeats.inn <- 1
# indep signif parameters
pthres <- 0.05
test <- "t.test"
method <- "none"
```

## Save up !!

```{r save}
# for cluster jobs, write out full combinations of double nested CV parameters
# - for outter loop, 1:(nfolds * nrepeats) denotes index for outter CV
# - for inner loop, 0 indicates running outter loop evaluation, 1:(nfolds.inn * nrepeats.inn) indicates index for inner CV
param <- expand.grid(as.character(xlist), # feature
                     as.character(ylist), # group
                     as.character(prlist), # predictor
                     as.integer(1:(nfolds * nrepeats)), # outter CV folds index for evaluation
                     as.integer(nfolds), 
                     as.integer(nrepeats), 
                     as.integer(0:(nfolds.inn * nrepeats.inn)), # inner CV folds for tuning predictor
                     as.integer(nfolds.inn), 
                     as.integer(nrepeats.inn), 
                     KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE)
write.table(param, file = '2_runPredict.txt', quote = FALSE, row.names = FALSE, col.names = FALSE, sep = ' ')
# preview of 2_runPredict.txt
str(param)

xlist.test <- grep("[.]and[.]|^genes", xlist, value = TRUE, invert = TRUE)
param <- expand.grid(as.character(xlist.test), # feature
                     as.character(ylist), # group
                     as.integer(1:(nfolds * nrepeats)), # outter CV folds index for evaluation
                     as.integer(nfolds), 
                     as.integer(nrepeats), 
                     as.numeric(pthres), 
                     as.character(test), 
                     as.character(method), 
                     KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE)
write.table(param, file = '4_indepsignif.txt', quote = FALSE, row.names = FALSE, col.names = FALSE, sep = ' ')
# preview of 4_indepsignif.txt
str(param)

xlist.nopen <- grep("genes", xlist, value = TRUE, invert = TRUE) # pre-selected features included in model
xlist.pen <- c("genes.vals") # select features while making prediction
xlist.fs <- as.vector(outer(xlist.nopen, xlist.pen, FUN = "paste", sep = ".fs.")) # see fsPred.R for details
prlist.fs <- c("predictorLogitLasso")
param <- expand.grid(as.character(xlist.fs), # feature
                     as.character(ylist), # group
                     as.character(prlist.fs), # predictor
                     as.integer(1:(nfolds * nrepeats)), # outter CV folds index for evaluation
                     as.integer(nfolds), 
                     as.integer(nrepeats), 
                     as.integer(0:(nfolds.inn * nrepeats.inn)), # inner CV folds for tuning predictor
                     as.integer(nfolds.inn), 
                     as.integer(nrepeats.inn), 
                     KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE)
write.table(param, file = '6_fsPredict.txt', quote = FALSE, row.names = FALSE, col.names = FALSE, sep = ' ')
# preview of 6_fsPredict.txt
str(param)

# save entire image to be loaded later
save(list = c(xlist, ylist, kmatlist), file = 'dat.RData')
```

# Session info

```{r session_info}
sessionInfo()
```
