# Results
# Yunlong Jiao, 10 March 2016

This script collects results from earlier run and illustrates with tables and plots.

```{r setup, message = FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 8, dev = c("png","pdf"), fig.keep = "high", fig.path = "result_figure/", cache.path = "result_cache/")
set.seed(70236562)
source("../../src/func.R")
library(ggplot2)
```

# Read in scores

```{r read_scores}
# read in score table
scores <- read.table("scores.txt", header = FALSE)
colnames(scores) <- c("x", "y", "predictor", "score", "value")
xlist <- unique(scores$x)
ylist <- unique(scores$y)
prlist <- unique(scores$predictor)
slist <- unique(scores$score)
# round values to 3 digits for ease of labeling
scores <- subset(scores, !is.na(scores$value))
scores$value <- round(scores$value, 3)
head(scores)
```

# Overview

We show the score values for different each feature types (with variance of boxplots across different methods in boxplot).

```{r overview}
# plot each grps in a separate figure
for (yname in ylist) {
  d <- subset(scores, scores$y == yname)
  p1 <- ggplot(d, aes(x = x, y = value)) + 
    geom_boxplot(aes(fill = x), alpha = 0.8) + 
    facet_wrap(~score) + 
    coord_cartesian(ylim = c(0, 1)) + 
    ggtitle(paste0("predicting ", yname)) + 
    theme(axis.text.x = element_blank())
  plot(p1)
}
```

# AUROC

As we see that class sizes are unbalanced so that we focus on AUROC to evaluate performance.

```{r auroc}
# focus only one type of scores
key <- "auroc"
# plot each grps in a separate figure
for (yname in ylist) {
  d <- subset(scores, scores$score == key & scores$y == yname)
  yrange <- c(min(d$value)*0.95, 1)
  p1 <- ggplot(d, aes(x = x, y = value)) + 
    geom_bar(aes(fill = x), stat = "identity", position = "dodge") + 
    geom_text(aes(label = value), vjust = 1.6, colour = "white", 
              position = position_dodge(0.9), size = 5) + 
    facet_wrap(~predictor) + 
    coord_cartesian(ylim = yrange) + 
    ggtitle(paste0(key, " for predicting ", yname)) + 
    theme(axis.text.x = element_blank())
  plot(p1)
}
```

# Session info

```{r session_info}
sessionInfo()
```
