# Feature selection results
# Yunlong Jiao, 10 June 2016

This script studies the potential of improving prediction by adding to pre-selected pathways a list of genes that provide othogonal information. Specifically a predictor is built on pathways features to make prediction while automating feature selection; this set of pre-selected pathway features are combined with genes to make a second/final model where the pre-selected pathways are always included in model but genes are being selected to improve the prediction.

```{r setup, message = FALSE}
knitr::opts_chunk$set(error = FALSE, fig.width = 12, fig.height = 8, dev = c("png","pdf"), fig.keep = "high", fig.path = "7_fsResults_figure/", cache.path = "7_fsResults_cache/")
set.seed(30430397)
source("../../src/func.R")
library(reshape2)
library(ggplot2)
```

First read in parameters!

```{r param}
# load in workspace
load("dat.RData")
rm(list = ls(pattern="[.]kmat$"))
# read in parameters
param <- read.table("6_fsPredict.txt", header = FALSE, row.names = NULL, col.names = c("xname", "yname", "prname", "i.fold", "nfolds", "nrepeats", "i.fold.inn", "nfolds.inn", "nrepeats.inn"))
# features
xlist <- unique(param$xname)
xlist <- grep("eff|path", xlist, value = TRUE) # only retain pathway values
xlist <- c(xlist.fs, # specific 2-step selected features
           unique(unlist(strsplit(xlist, "[.]fs[.]"))), # individual features
           gsub("vals[.]fs", "and", xlist), # "and"-type features
           "fun.vals", "go.vals", "mini.genes.vals", "other.genes.vals") # two more restricted set of gene-related vals
xlist
# feature types # the following order is fairly important
xlist.type <- c("func-wise", "func-wise", 
                "path-wise", "path-wise", 
                "gene-wise", "gene-wise", "gene-wise", 
                "mix-wise", "mix-wise", 
                "fs-wise", "fs-wise")
names(xlist.type) <- c("fun.vals", "go.vals", # functionality features
                       "eff.vals", "path.vals", # pathway features
                       "mini.genes.vals", "other.genes.vals", "genes.vals", # gene features
                       "eff.and.genes.vals", "path.and.genes.vals", 
                       "eff.vals.fs.genes.vals", "path.vals.fs.genes.vals") # mixed type
xlist.vline <- c(2.5, 4.5, 7.5, 9.5) # cut out types
stopifnot(length(setdiff(xlist, names(xlist.type))) == 0)
xlist.type
# groups
ylist <- unique(param$yname)
ylist
# predictors
prlist <- unique(param$prname)
prlist
stopifnot(all(prlist == "predictorLogitLasso"))
# (outter) `nfolds`-fold CV repeated `nrepeats` times for evaluation
nfolds <- unique(param$nfolds)
stopifnot(length(nfolds) == 1)
nfolds
nrepeats <- unique(param$nrepeats)
stopifnot(length(nrepeats) == 1)
nrepeats
# inner `nfolds.inn`-fold CV repeated `nrepeats.inn` times for tuning predictors
nfolds.inn <- unique(param$nfolds.inn)
stopifnot(length(nfolds.inn) == 1)
nfolds.inn
nrepeats.inn <- unique(param$nrepeats.inn)
stopifnot(length(nrepeats.inn) == 1)
nrepeats.inn
# evaluation measures
slist <- c("acc","fpr","tpr","ppv","fval", "auroc")
slist
slist.prefer.large.score <- c(TRUE, FALSE, TRUE, TRUE, TRUE, TRUE)
names(slist.prefer.large.score) <- slist
slist.prefer.large.score
```


## Performance with logitLasso only

In this secion, we fix the predictor to be logitLasso and compare model performances of different feature types wrt various criteria.

```{r perfLogitLasso_prepare, message=FALSE}
prname <- "predictorLogitLasso"
scores <- list()
for (yname in ylist) {
  for (xname in xlist) {
    message(yname, '\t', xname)
    res.files <- list.files(path = 'Robj', 
                            pattern = paste('^ivres', xname, yname, prname, 
                                            '[[:digit:]]+', nfolds, nrepeats, 
                                            0, nfolds.inn, nrepeats.inn, 
                                            sep = '_'), 
                            full.names = TRUE)
    res <- lapply(res.files, function(f) get(load(f)))
    names(res) <- paste0("rep",1:length(res))
    # reduce to only scores and melt
    res <- lapply(res, '[', slist)
    res <- melt(res)
    colnames(res)[colnames(res) == "L1"] <- "rep"
    colnames(res)[colnames(res) == "L2"] <- "score"
    
    # record score values
    objname <- paste('res', xname, yname, prname, nfolds, nrepeats, sep = '_')
    scores[[objname]] <- data.frame(y = yname, 
                                    x = xname, 
                                    type = xlist.type[xname], 
                                    predictor = prname, 
                                    res, 
                                    row.names = NULL)
  }
}
scores <- do.call('rbind', scores)
rownames(scores) <- seq(nrow(scores))
head(scores)
```

```{r perfLogitLasso}
# plot each grps in a separate figure
for (yname in ylist) {
  d <- subset(scores, scores$y == yname)
  # order feature by types
  d$x <- factor(d$x, levels = names(xlist.type), ordered = TRUE)
  # change score values by "fpr" to that by "spec"
  d$value[d$score == "fpr"] <- 1 - d$value[d$score == "fpr"]
  d$score[d$score == "fpr"] <- "spec"
  # change score name "tpr" to "sens"
  d$score[d$score == "tpr"] <- "sens"
  p1 <- ggplot(d, aes(x = x, y = value)) + 
    geom_boxplot(aes(fill = x), alpha = 0.8) + 
    geom_vline(xintercept = xlist.vline, color = "grey", size = 2) + 
    facet_wrap(~score, scales = "free") + 
    ggtitle(paste0("predicting ", yname, " using ", prname)) + 
    theme(axis.text.x = element_blank())
  plot(p1)
}
```


## Feature selection

In this section, with the predictor built following a specific 2-step feature selection, we have found **genes that provide orthogonal information**, compensating (and improving) the prediction from pathway features for each `yname` in `r ylist`.

```{r featselect_prepare, message=FALSE}
xname <- "path.vals.fs.genes.vals" # do not alter
xname.list <- c("mini.genes.vals", "other.genes.vals", "path.vals") # do not alter
featlist.long <- lapply(xname.list, function(u){
  colnames(get(u))
})
names(featlist.long) <- xname.list
n.featlist.long <- sapply(featlist.long, length)
# show each grp separately
for (yname in ylist) {
  for (prname in prlist) {
    message(yname, '\t', prname)
    res.files <- list.files(path = 'Robj', 
                            pattern = paste('^ivres', xname, yname, prname, 
                                            '[[:digit:]]+', nfolds, nrepeats, 
                                            0, nfolds.inn, nrepeats.inn, 
                                            sep = '_'), 
                            full.names = TRUE)
    res <- lapply(res.files, function(f) get(load(f)))
    names(res) <- paste0("rep",1:length(res))
    
    # featselect
    featlist.short <- lapply(res, '[[', 'featlist.short')
    featlist.short <- lapply(featlist.long, function(u){
      lapply(featlist.short, function(v){
        intersect(u,v)
      })
    })
    assign(paste(yname,prname,"featlist.short",sep="_"), featlist.short)
    rm(res)
  }
}
```

For each `yname` in `r ylist`, we have several plots to illustrate the results
- Boxplot on the number of selected features from each feature type in `r xname.list`, wrt different algorithms, where variance shows at each CV run.
- Histogram of how many times one feature are selected in total over all `r nfolds*nrepeats` CV runs, specific to each type and each algorithm.

```{r featselect}
# show each grp separately
for (yname in ylist) {
  cat('\n---> \t Specific 2-step feature selection \t <---\n')
  cat('\n---------> \t for ', yname, ' \t <---------\n')
  # featselect
  featlist.short <- lapply(prlist, function(prname){
    get(paste(yname,prname,"featlist.short",sep="_"))
  })
  names(featlist.short) <- prlist
  # featlist.short is 3-layer list prlist -> xname.list -> CV repeats -> a character vector of feature names
  # number of features are selected at each CV run
  featlist.freq <- lapply(featlist.short, function(u) lapply(u, lapply, length))
  featlist.freq <- melt(featlist.freq, level = 1, value.name = "value")
  colnames(featlist.freq)[colnames(featlist.freq)=="L1"] <- "prname"
  colnames(featlist.freq)[colnames(featlist.freq)=="L2"] <- "xname"
  colnames(featlist.freq)[colnames(featlist.freq)=="L3"] <- "rep"
  p1 <- ggplot(featlist.freq, aes(x = xname, y = value)) + 
    geom_boxplot(aes(fill = xname), alpha = 0.8) + 
    facet_wrap(~prname, scales = "free") + 
    scale_x_discrete(name = "x") + 
    scale_y_continuous(name = "Number of selected features") + 
    ggtitle(paste0("How many features are selected at each run")) + 
    theme(axis.text.x = element_blank(), legend.title = element_blank(), legend.position = "bottom")
  plot(p1)
  
  # proportion across group
  n.featlist.short <- tapply(featlist.freq$value, list(featlist.freq$rep, featlist.freq$prname), sum)
  cat('\nTotal number of features selected at each run (averaged over each CV run)\n')
  cat('\nNOTE THAT GENES PROVIDE ORTHOGONAL INFO TO PATHWAYS\n')
  print(colMeans(n.featlist.short))
  
  # proportion within group
  cat('\nTotal number of features within each type\n')
  cat('\nNOTE THAT GENES PROVIDE ORTHOGONAL INFO TO PATHWAYS\n')
  print(n.featlist.long)
  
  # show top 10 most selected features in each type by predictor
  featlist.tab <- lapply(featlist.short, function(u) lapply(u, function(v) 
    sort(table(unlist(v)), decreasing = TRUE)
  ))
  cat('\nPreview of top 10 most often selected features in each type\n')
  cat('\nNOTE THAT GENES PROVIDE ORTHOGONAL INFO TO PATHWAYS\n')
  print(lapply(featlist.tab, function(u) lapply(u, head, n=10)))
  
  # count number of times of features from each type being selected over CV runs
  featlist.count <- melt(featlist.tab, level = 1, value.name = "value")
  colnames(featlist.count)[colnames(featlist.count)=="L1"] <- "prname"
  colnames(featlist.count)[colnames(featlist.count)=="L2"] <- "xname"
  colnames(featlist.count)[colnames(featlist.count)=="Var1"] <- "featname"
  featlist.count <- dcast(featlist.count, prname+xname~value, fun.aggregate = length)
  featlist.count <- melt(featlist.count, id.vars = c("prname","xname"))
  p1 <- ggplot(featlist.count, aes(x = variable, y = value)) + 
    geom_bar(aes(fill = xname), stat = "identity", position = "dodge", alpha = 0.8) + 
    geom_text(aes(label = value, group = xname), size = 2, angle = 45, colour = "black", position = position_dodge(0.9), vjust = 0.9) + 
    facet_wrap(~prname, scales = "free", ncol = 1) + 
    scale_x_discrete(name = paste0("selected times (max ", nfolds*nrepeats, ")")) + 
    scale_y_continuous(name = "count") + 
    ggtitle(paste0("How many times features are selected over all ", nfolds*nrepeats, " CV runs")) + 
    theme(legend.title = element_blank(), legend.position = "bottom")
  plot(p1)
}
```



## Session info

```{r session_info}
sessionInfo()
```
